---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/bain)](https://cran.r-project.org/package=bain)
[![Build Status](https://travis-ci.org/cjvanlissa/bain.svg?branch=master)](https://travis-ci.org/cjvanlissa/bain)
[![](https://cranlogs.r-pkg.org/badges/bain)](https://cran.r-project.org/package=bain)
<!--[![Codecov test coverage](https://codecov.io/gh/cjvanlissa/bain/branch/master/graph/badge.svg)](https://codecov.io/gh/cjvanlissa/bain?branch=master)-->

# bain

Bain stands for Bayesian informative hypothesis evaluation. It computes Bayes factors for informative hypotheses in a wide variety of statistical models. Just run your analysis as usual, and then apply bain to the output. A tutorial is available at [DOI:10.31234/osf.io/v3shc](https://psyarxiv.com/v3shc/).

## Installation

Install the latest release version of `bain` from CRAN:

```{r cran-installation, eval = FALSE}
install.packages("bain")
```

You can also install the latest development version of `bain` from GitHub. This requires a working toolchain, to compile the Fortran source code. [Step 3 in this tutorial](https://cjvanlissa.github.io/worcs/articles/setup.html) explains how to set up the toolchain. Then, run:

```{r github-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("cjvanlissa/bain")
```

## Workflow

Add bain to your existing R workflow, and obtain Bayes factors for your familiar R analyses! Bain is compatible with the pipe operator. Here is an example for testing an informative hypothesis about mean differences in an ANOVA:

```{r example, eval = TRUE, message = FALSE}
# Load bain
library(bain)
# dplyr to access the %>% operator
library(dplyr)
# Iris as example data
iris %>%
  # Select outcome and predictor variables
  select(Sepal.Length, Species) %>%      
  # Add -1 to the formula to estimate group means, as in ANOVA
  lm(Sepal.Length ~ -1 + Species, .) %>% 
  bain("Speciessetosa < Speciesversicolor = Speciesvirginica;
       Speciessetosa < Speciesversicolor < Speciesvirginica")
```

## Documentation

Every user-facing function in the package is documented, and the documentation can be accessed by running `?function_name` in the R console, e.g., `?bain`.

Moreover, you can read the *Introduction to bain* vignette by running `vignette("Introduction_to_bain", package = "bain")`

## Citing bain

You can cite the R-package with the following citation:

> Gu, X., Hoijtink, H., Mulder, J., & van Lissa, C. (2019). bain: Bayes factors for informative hypotheses. (Version 0.2.3) [R package]. https://CRAN.R-project.org/package=bain

## Contributing and Contact Information

If you have ideas, please get involved. You can contribute by opening an issue on GitHub, or sending a pull request with proposed features.

* File a GitHub issue [here](https://github.com/cjvanlissa/bain)
* Make a pull request [here](https://github.com/cjvanlissa/bain/pulls)

By participating in this project, you agree to abide by the [Contributor Code of Conduct v2.0](https://www.contributor-covenant.org/).
