---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/bain)](https://cran.r-project.org/package=bain)
[![Build Status](https://travis-ci.org/cjvanlissa/bain.svg?branch=master)](https://travis-ci.org/cjvanlissa/bain)
[![](https://cranlogs.r-pkg.org/badges/tidyLPA)](https://cran.r-project.org/package=tidyLPA)

# bain

Bain stands for Bayesian informative hypothesis evaluation. It computes Bayes factors for informative hypotheses in a wide variety of statistical models. Just run your analysis as usual, and then apply bain to the output. A tutorial is available at [DOI:10.31234/osf.io/v3shc](https://psyarxiv.com/v3shc/).

## Installation

Install `bain` from CRAN:

```{r cran-installation, eval = FALSE}
install.packages("bain")
```

## Workflow

Add bain to your existing R workflow, and obtain Bayes factors for your familiar R analyses! Bain is compatible with the pipe operator. Here is an example for testing an informative hypothesis about mean differences in an ANOVA:

```{r example, eval = TRUE, message = FALSE}
# Load bain
library(bain)
# dplyr to access the %>% operator
library(dplyr)
# Iris as example data
iris %>%
  # Select outcome and predictor variables
  select(Sepal.Length, Species) %>%      
  # Add -1 to the formula to estimate group means, as in ANOVA
  lm(Sepal.Length ~ -1 + Species, .) %>% 
  bain("Speciessetosa < Speciesversicolor = Speciesvirginica;
       Speciessetosa < Speciesversicolor < Speciesvirginica")
```
